\section{Discussion}
% and Concluding remarks}
\label{sec:disc}

\subsection{Extending the protocol to universally commutative commands}

% Generalized Paxos vs Paxos spec.
The Generalized Paxos specification dictates that the only commands
whose relative order can be reversed at different learners are those
that commute with each other. This restriction has the advantage of
allowing for running a replicated state machine on top of the
Generalized Paxos protocol, while still ensuring that the state
machine behaves in a way that is indistinguishable from running it on
top of the original Paxos protocol.

% Extension to ops that commute with all
However, the downside of this use of commutative operations in the
context of Generalized Paxos is that this commutativity check is done
at runtime, and, if two non-commutative operations are issued
concurrently in the same ballot, then we must fall back to the slower
classic Paxos protocol.

This raises the possibility of extending the protocol to handle
commands that are universally commutative, i.e., commute with every
other command. For these commands, it is known before executing them
that they will not generate any conflicts, and therefore it is not
necessary to check them against concurrently executing commands.  This
allows us to optimize the protocol by decreasing the number of phase
$2b$ messages required to learn to a smaller $f+1$ quorum. Since, by
definition, these sequences are guaranteed to never produce conflicts,
the $N-f$ quorum is not required to prevent learners from learning
conflicting sequences. Instead, a quorum of $f+1$ is sufficient for
the learner to be sure that the proposed sequence was proposed by a
correct proposer. (Furthermore, for read-only commands that do not
change the state of the state machine, it is possible to run these
against one single learner in the crash model, since there is no
requirement to withstand $f$ faults for these commands.)
This optimization is particularly useful in the context of 
geo-replicated systems, since it can be significantly faster
to wait for the $f+1$st message instead of the $N-f$th one.

% Weakly consistent replication - also diff order but ok to have different v

\subsection{Generalized Paxos and weak consistency}


%The Byzantine Generalized Paxos protocol tackles two challenges in two different avenues of research, fault tolerance and relaxed consistency models. By specifying the generalized consensus problem,
The key distinguishing feature of the specification of Generalized
Paxos~\cite{Lamport2005} is allowing learners to learn concurrent
proposals in a different order, when the proposals commute. This idea
is closely related to the work on weaker consistency models like eventual or
causal consistency~\cite{Ahamad1995}, or consistency models that mix
strong and weak consistency levels like RedBlue~\cite{Li2012}, which attempt
to decrease the cost of executing operations by reducing coordination
requirements between replicas. 

The link between the two models becomes clearer after we introduce the
notion of universally commutative commands from the previous subsection.
In the case of weakly consistent replication (or multi-level replication),
weakly consistent requests can be executed as if they were universally
commutative, even if in practice that may not be the case. E.g., checking 
the balance of a bank account and making a deposit do not commute since
the output of the former depends on their relative order. However,
some systems prefer to run both as weakly consistent operations, even
though it may cause executions that are not explained by a sequential
execution of all commands, since the semantics are still acceptable given
that the final state that is reached is the same and no invariants 
of the application are violated~\cite{Li2012}.



% Extension to diff replica grps


\subsection{Handling faults in the fast case}

A result that was stated in the original Generalized Paxos
paper~\cite{Lamport2005} is that to tolerate $f$ crash faults and
allow for fast ballots whenever there are up to $e$ crash faults, the
total system size $N$ must uphold two conditions:
\begin{itemize}
	\item $N > 2f$
	\item $N > 2e+f$
\end{itemize} 

Additionally, the fast and classic quorums are of size $N-e$ and $N-f$, respectively, which means there is a dependence relationship between these three concepts: number of tolerated faults, system size, and quorum size. As such, if we don't mind keeping $e$ small, then the fast ballot quorum will be a larger fraction of the total system size, which itself will be smaller. Conversely, if we wish to allow fast ballots to progress in the presence of the maximum number of faults, we can set $e=f$, which means that $N >3f$ and $N-e=N-f$. \par 
An interesting observation stemming from our results is that, since Byzantine fault tolerance also requires a total system size of $N>3f$ and a quorum size of $2f+1$, we are able to amortize the cost of both features, i.e., we are able to tolerate the maximum number of faults for fast execution while not increasing the overall replication factor.

\subsection{Comparison to Fast Byzantine Consensus}
Previous fast protocols that solve consensus in a Byzantine settings include Fast Byzantine Paxos (FaB)~\cite{Martin2006}. FaB solves consensus in two communication steps in the common case while requiring $5f+1$ acceptors to ensure safety and liveness. The Parameterized FaB Paxos protocol generalizes FaB that requires $3f+2t+1$ to solve consensus while tolerating up to $f$ faults and is fast as long as up to $t$ acceptors fail. Therefore, FaB Paxos is a special case of Parameterized FaB Paxos where $t=f$. \par
FaB Paxos ensures safety as long as $N>3f$ but requires $N
>5f$ to ensure liveness. FaB is limited by a lower bound of $N > 5f$ acceptors in order to guarantee both safety and liveness. However, BGP allows several commands to be learned in the ballot and even no command at all. \par
One potentially dangerous scenario occurs when the leader is faulty. In this case, the leader could split the acceptors votes or run a fast ballot in some and a classic ballot in others. In this scenario, since learners require $N-f$ votes for any given sequence and non-commutative sequences aren't considered equivalent, at most one sequence can be learned by correct learners. This also applies to the case where two non-commutative sequences are proposed in a fast ballot. It's possible for the votes to be split in a way such that no sequence is learned. In this case, for the remainder of the ballot, since every possible subsequent sequence at each acceptor will be an extension of the previous non-commutative sequences, no new sequences will be learned. This prevents safety from being violated but also precludes liveness until a new classic ballot is initiated by a correct leader. Once a correct leader gathers the previous votes from the acceptors, it will propose a serialization of previous sequences that solves the conflict caused by the non-commutative sequences.

\section{Conclusion}
\label{sec:conc}
In this paper, we presented a simplified description of the Generalized Paxos protocol and specification, which is meant to pave the way for new avenues of research in this area. In addition, we present a Byzantine fault tolerant version of the protocol, and we prove the correctness of this protocol. In the future, we would like to implement this protocol and fully evaluate it, in addition to gaining a better understanding of its practical applicability.
