\subsection{Discussion}
% and Concluding remarks}
\label{sec:disc}


\subsubsection{Handling faults in the fast case}
A result that was stated in the original Generalized Paxos
paper~\cite{Lamport2005} is that to tolerate $f$ crash faults and
allow for fast ballots whenever there are up to $e$ crash faults, the
total system size $N$ must uphold two conditions:
$N > 2f$ and $N > 2e+f$.

Additionally, the fast and classic quorums are of size $N-e$ and $N-f$, respectively, which means there is a dependence relationship between these three concepts: number of tolerated faults, system size, and quorum size. As such, if we don't mind keeping $e$ small, then the fast ballot quorum will be a larger fraction of the total system size, which itself will be smaller. Conversely, if we wish to allow fast ballots to progress in the presence of the maximum number of faults, we can set $e=f$, which means that $N >3f$ and $N-e=N-f$. 


An interesting observation stemming from our results is that, since Byzantine fault tolerance also requires a total system size of $N>3f$ and a quorum size of $2f+1$, we are able to amortize the cost of both features, i.e., we are able to tolerate the maximum number of faults for fast execution while not increasing the overall replication factor.
{\color{red}Given the subsection stating that our relaxed problem can forgo of 3f+1 while allowing fast learning, this needs to be restated from "amortize the costs" to "since we eliminated the previous cost of fast learning, we can now support Byzantine fault tolerance for the same cost that Fast/Generalized Paxos require to tolerate crash faults}

\subsubsection{Fast Byzantine Paxos Comparison}
In comparison to FaB Paxos, our protocol, Byzantine Generalized Paxos (BGP), requires a lower number of acceptors than what is stipulated by FaB's lower bound. However, this does not constitute a violation of the result since BGP does not guarantee a 2-step execution in the Byzantine scenario. Instead, BGP only provides a two communication step latency when proposed sequences are commutative with any other concurrently proposed sequence. In other words, Byzantine Generalized Paxos leverages a weaker performance guarantee to decrease the requirements regarding the minimum number of processes. In particular, a proposed sequence may not gather enough votes to be learned in the ballot in which it is proposed, either due to Byzantine behaviour or contention between non-commutative commands. However, any sequence is guaranteed to eventually be learned in a way such that non-commutative commands are totally ordered at any correct learner.